---
title: "Weekly Sales Report"
author: "Waliu Ayuba"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(DBI)
library(RMySQL)
library(ggplot2)

# Connect to MySQL Database
con <- dbConnect(RMySQL::MySQL(), 
                 dbname = "sql5699606", 
                 host = "sql5.freemysqlhosting.net",
                 port = 3306, 
                 user = "sql5699606", 
                 password = "GY1IlvutbB")


```

## Analytical Query I: Top Five Sales Reps by Year
```{r echo=F, warning=FALSE}
query <- "
SELECT year, rep_name, SUM(total_amount_sold) as total_sales
FROM rep_facts
GROUP BY year, rep_name
ORDER BY year, total_sales DESC
LIMIT 5"
results <- dbGetQuery(con, query)
knitr::kable(results, caption = 'Top Five Sales Representatives by Year')

```



## Analytical Query II: Total Sales Per Product Per Quarter
```{r echo=FALSE, warning=FALSE}
query <- "
SELECT year, quarter, product_name, SUM(total_amount_sold) as total_sales
FROM product_facts
GROUP BY year, quarter, product_name"
results <- dbGetQuery(con, query)
knitr::kable(results, caption = 'Total Sales Per Product Per Quarter')


```


## Analytical III: Units Sold Per Product Per Region
```{r echo=FALSE, warning=FALSE}
query <- "SELECT product_name, region, SUM(total_units_sold) as units_sold FROM product_facts GROUP BY product_name, region"
results <- dbGetQuery(con, query)
ggplot(results, aes(x=region, y=units_sold, group=product_name, color=product_name)) + 
    geom_line() + 
    labs(title="Units Sold Per Product Per Region", x="Region", y="Units Sold")

```

Explanation:
The image displays a line graph titled "Units Sold Per Product Per Region." The graph plots the units sold of different pharmaceutical products across three regions: Brazil, Germany, and the USA. Each product is represented by a unique color, with the product names listed in the legend on the right.
From left to right, three points on the x-axis correspond to the three regions. The y-axis measures the number of units sold, ranging from 1 million to a little over 3 million units. All products seem to follow a similar trend:

- Fewer units sold in Brazil

- A significant drop in units sold in Germany

- A sharp increase in units sold in the USA

The lines connecting the points show the relative differences in units sold per product per region. Notably, there's a dramatic increase in the number of products in the USA, suggesting a larger market size or higher demand in the USA compared to Brazil and Germany. Germany shows the lowest number of units sold among the three regions for all products. 


## Analytical Query IV: Average Sales Per Sales Rep Over the Years

```{r average_sales_rep_year, echo=FALSE, warning=FALSE, }
query <- "
SELECT year, rep_name, AVG(average_amount_sold) as avg_sales
FROM rep_facts
GROUP BY year, rep_name
ORDER BY year, rep_name"

results <- dbGetQuery(con, query)

# Convert the year to an integer if it is not already
results$year <- as.integer(results$year)

# Check if results are empty
if(nrow(results) == 0) {
  cat("No data available for plotting.")
} else {
  ggplot(results, aes(x=as.factor(year), y=avg_sales, color=rep_name, group=rep_name)) +
    geom_line() +
    geom_point() +
    labs(title="Average Sales Per Sales Rep Over the Years", x="Year", y="Average Sales ($)", color="Sales Rep") +
    theme_minimal()
}


```

The line graph, titled "Average Sales Per Sales Rep Over the Years," is a crucial tool for understanding the sales performance of our representatives. It represents the average sales in dollars achieved by different sales representatives over a period spanning from 2020 to 2022.
As indicated in the legend on the right, each sales representative is represented by a different color. The x-axis represents time, with a fractional part of the year (e.g., 2020 indicates the midpoint of 2020), while the y-axis represents the average sales in dollars.
Let's delve into the insights this graph offers:

- The sales performance of each representative varies significantly over the years.

- Veronika Sixt (light blue line) shows a dramatic increase in average sales towards 2022, suggesting a significant performance improvement or a successful sales initiative.

- Conversely, Walison da Silva (magenta line) started with higher average sales in 2020 but showed a decreasing trend over the next two years.

- Aneeta Kappoorthy (red line) shows some fluctuations but ends slightly higher in 2022 than where they started in 2020.

- Other representatives show various trends of increases and decreases over the years with no consistent pattern across the group.
