---
title: "Query Tabular Data with SQL"
author: "Waliu Ayuba"
date: "Spring 2024"
output:
  html_document:
    df_print: paged
---

```{r }
library(readr)
library(sqldf)
library(dplyr)

#URLs of the datasets
url_jan_mar <- "https://s3.us-east-2.amazonaws.com/artificium.us/assignments/80.xml/a-80-305/gen-xml/synthsalestxns-Jan2Mar.csv"
url_sep_oct <- "https://s3.us-east-2.amazonaws.com/artificium.us/assignments/80.xml/a-80-305/gen-xml/synthsalestxns-Sep2Oct.csv"
url_nov_dec <- "https://s3.us-east-2.amazonaws.com/artificium.us/assignments/80.xml/a-80-305/gen-xml/synthsalestxns-Nov2Dec.csv"

#Load the datasets
data_jan_mar <- read_csv(url_jan_mar)
data_sep_oct <- read_csv(url_sep_oct)
data_nov_dec <- read_csv(url_nov_dec)

#Combine the datasets
all_data <- bind_rows(data_jan_mar, data_sep_oct, data_nov_dec)


```


## Total visits and Unique Restaurants
```{r}
# SQL query for total visits and unique restaurants
total_unique <- sqldf("SELECT COUNT(*) AS Total_Visits, COUNT(DISTINCT restaurant) AS Unique_Restaurants FROM all_data")

total_unique

```


## Time Range of the Data
```{r}

#Convert 'amount' to nemric
all_data$amount <- as.numeric(gsub("\\$", "", all_data$amount))

# Convert 'date' column to Date format
all_data$date <- as.Date(all_data$date, format="%m/%d/%y")

# SQL query for time range(Extract Year)
date_range <- sqldf("SELECT MIN(date) AS Start_Date, MAX(date) AS End_Date FROM all_data")

date_range$Start_Year <- format(as.Date(date_range$Start_Date), "%Y")
date_range$End_Year <- format(as.Date(date_range$End_Date), "%Y")

date_range$Start_Year
date_range$End_Year
```


## Average Amount Spent
```{r}
# Convert 'amount' to numeric
all_data$amount <- as.numeric(gsub("\\$", "", all_data$amount))

# SQL query for average amount and standard deviation
avg_std_amount <- sqldf("SELECT AVG(amount) AS Average_Amount, STDEV(amount) AS Std_Dev_Amount FROM all_data")

avg_std_amount

```


## Restaurant with Most Visits and Largest Revenue
```{r}
# Most visits
most_visits <- sqldf("SELECT restaurant, COUNT(*) AS Visits FROM all_data GROUP BY restaurant ORDER BY Visits DESC LIMIT 1")

most_visits

# Largest revenue
largest_revenue <- sqldf("SELECT restaurant, SUM(amount) AS Total_Revenue FROM all_data GROUP BY restaurant ORDER BY Total_Revenue DESC LIMIT 1")

largest_revenue
```

## Average Party Size and Percentage of Parties Consuming Wine
```{r}


# Average party size and wine consumption percentage
party_wine_stats <- sqldf("SELECT AVG(guests) AS Average_Party_Size, AVG(CASE WHEN wine = 'Yes' THEN 1 ELSE 0 END) * 100 AS Wine_Consumption_Percentage FROM all_data")

party_wine_stats
```

### We recorded a total number of `r total_unique$Total_Visits` visits across `r total_unique$Unique_Restaurants` different restaurants from `r date_range$Start_Year` to `r date_range$End_Year`. The average amount spent for a visit across all restaurants is `r round(avg_std_amount$Average_Amount, 1)` (σ = `r round(avg_std_amount$Std_Dev_Amount, 1)`). The most visits were at `r most_visits$restaurant` while the largest revenue across all years was at `r largest_revenue$restaurant`. The average size of a party was `r round(party_wine_stats$Average_Party_Size, 0)` and `r round(party_wine_stats$Wine_Consumption_Percentage, 1)`% of all parties consumed wine.


## Summary of Revenue
```{r}

# Summary of revenue by restaurant
revenue_summary <- sqldf("SELECT restaurant, COUNT(*) AS Total_Visits, AVG(amount) AS Average_Revenue FROM all_data GROUP BY restaurant")

revenue_summary

```


