---
title: "R Notebook:Query a Database with SQL"
author: "By Waliu Ayuba"
output:
  html_document:
    df_print: paged
---



```{r}
library(RSQLite)
dbfile <-"MediaDB.db"
dbcon <- dbConnect(SQLite(), dbfile)
```


## 1.What are the last name, first name, title, and hire date of all employees, sorted by last name?
```{r}
query <- "SELECT Lastname, FirstName, Title, HireDate  
FROM employees
ORDER BY LastName;"
result <- dbGetQuery(dbcon, query)
print(result)
```



## 2. What are the names of each genre and the number of distinct albums in that genre? 
```{r}
query <- "SELECT g.Name AS GenreName, COUNT(DISTINCT a.AlbumId) AS NumberOfDistinctAlbums
FROM genres g
JOIN tracks t ON g.GenreId = t.GenreId
JOIN albums a ON t.AlbumId = a.AlbumId
GROUP BY g.Name
ORDER BY NumberOfDistinctAlbums DESC;"

result <- dbGetQuery(dbcon, query)
print(result)
```


## 3.What are the distinct media types?
```{r}
query <- "SELECT DISTINCT Name
          FROM media_types;"
result <- dbGetQuery(dbcon, query)
print(result)
```

## 4. List the number of tracks by genre, ordered from most to least, restricted to those genres for which there are at least four tracks.
```{r}
query <- "SELECT g.Name AS GenreName, COUNT(t.TrackId) AS NumberOfTrack
          FROM genres g
          JOIN tracks t ON g.GenreId = t.GenreId
          GROUP BY g.Name
          HAVING COUNT(t.TrackId) >= 4
          ORDER BY NumberOfTrack DESC"
result <- dbGetQuery(dbcon, query)
print(result)
```


## 5. What is the last name and first name of the youngest employee?
```{r}
query <- "SELECT LastName, FirstName
          FROM employees
          ORDER BY STRFTIME('%Y-%m', BirthDate) DESC
          LIMIT 1;"

result <- dbGetQuery(dbcon, query)
print(result)
```

## 6. What are the total number of unique purchases for each Brazilian state (e.g., MG, SP, etc.), ordered alphabetically by state, and in which there were at least two purchases?

```{r}
query <- "SELECT c.State, COUNT(DISTINCT i.InvoiceId) AS TotalPurchases
          FROM customers c
          JOIN invoices i ON c.CustomerId = i.CustomerId
          WHERE c.Country = 'Brazil'
          GROUP BY c.State
          HAVING COUNT(DISTINCT i.InvoiceId) >= 2
          ORDER BY c.State ASC;"

result <- dbGetQuery(dbcon, query)
print(result)
```

## 7. How many employees never worked with a single customer, i.e., they appear in the employees table but not in the customers table?
```{r}
query <- "SELECT COUNT(*) AS NumberOfEmployeesWithoutCustomers
          FROM employees e
          LEFT JOIN customers c ON e.employeeId = c.SupportRepId
          WHERE c.SupportRepId IS NULL;"

result <- dbGetQuery(dbcon, query)
print(result)
```

## 8. How many albums contain the substring "Classic" but not "Hard" in the album title?

```{r}

query <- "SELECT COUNT(*)
          FROM albums
          WHERE Title Like '%Classic%' AND Title NOT LIKE '%Hard%';"

result <- dbGetQuery(dbcon, query)
print(result)
```
## 9. What is the total number of distinct genres?

```{r}
query <- "SELECT COUNT(DISTINCT GenreId) AS TotalNumberOfDistinctGenres
          FROM genres;"

result <- dbGetQuery(dbcon, query)
print(result)
```


## 10. What is the total "time" of all tracks per media type in hours, i.e., add up the "Milliseconds" column in tracks? Display the media type and the time in hours rounded to one decimals. Only consider media types with at least one track.


```{r}

query <- "SELECT mt.Name AS MediaType,
          ROUND(SUM(tr.Milliseconds) / 3600000.0,1) AS TotalTimeInHours
          FROM media_types mt
          JOIN tracks tr ON mt.MediaTypeId = tr.MediaTypeId
          GROUP BY mt.Name
          HAVING COUNT(tr.TrackId) > 0;"

result <- dbGetQuery(dbcon, query)
print(result)
```


## 11. What is the title and artist name of the longest album, in terms of minutes of playtime of all tracks on that albums?

```{r}
query <- "SELECT al.Title AS AlbumTitle, ar.Name AS ArtistName, 
          SUM(tr.Milliseconds) / 60000.0 AS TotalMinutes
          FROM albums al
          JOIN tracks tr ON al.AlbumId = tr.AlbumId
          JOIN  artists ar ON al.ArtistId = ar.ArtistId
          GROUP BY al.AlbumId, al.Title, ar.Name
          ORDER BY TotalMinutes DESC
          LIMIT 1;"
result <- dbGetQuery(dbcon, query)
print(result)
```


## 12. Which genre has the least number of tracks?

```{r}
query <- "SELECT g.Name AS GenreName, COUNT(t.TrackId) AS NumberOfTracks
          FROM genres g
          JOIN tracks t ON g.GenreId = t.GenreId
          GROUP BY g.GenreId
          ORDER BY NumberOfTracks ASC
          LIMIT 1;
          "
result <- dbGetQuery(dbcon, query)
print(result)
```

## 13. Which media format has the fewest tracks?

```{r}
query <- "SELECT mt.Name AS MediaType, COUNT(t.TrackId) AS NumberOfTracks
          FROM media_types mt
          JOIN tracks t ON mt.MediaTypeId = t.MediaTypeId
          GROUP BY mt.MediaTypeId
          ORDER BY NumberOfTracks ASC
          LIMIT 1;"

result <- dbGetQuery(dbcon, query)
print(result)
```


## 14. For each country, what is the total number of unique customers and the total spent?

```{r}
query <- "SELECT c.Country, COUNT(DISTINCT c.CustomerId) AS TotalUniqueCustomers, 
          SUM(i.Total) AS TotalSpent
          FROM customers c
          JOIN invoices i ON c.CustomerId = i.CustomerId
          GROUP BY c.Country
          ORDER BY c.Country"

result <- dbGetQuery(dbcon, query)
print(result)
```


## 15. What is the total amount spent by all customers who do not live in South America?

```{r}
query <- "SELECT SUM(i.Total) AS TotalSpent
          FROM customers c
          JOIN invoices i ON c.CustomerId = i.CustomerId
          WHERE c.Country NOT IN('Argentina', 'Bolivia', 'Brazil', 'Chile', 'Colombia', 
          'Ecuador', 'Guyana', 'Paraguay', 'Peru', 'Suriname', 'Uruguay', 'Venezuela')"
result <- dbGetQuery(dbcon, query)
print(result)
```

## 16. What is the difference in spending on any type of MPEG vs any type of AAC track?

```{r}
query <- "SELECT SUM(CASE WHEN mt.Name LIKE '%MPEG%' THEN ii.UnitPrice * ii.Quantity ELSE 0 END) AS TotalMPEG,
          SUM(CASE WHEN mt.Name LIKE '%AAC%' THEN ii.UnitPrice * ii.Quantity ELSE 0 END) AS TotalAAC,
          ABS(SUM(CASE WHEN mt.Name LIKE '%MPEG%' THEN ii.UnitPrice * ii.Quantity ELSE
          0 END) - SUM(CASE WHEN mt.Name LIKE '%AAC%' THEN ii.UnitPrice * ii.Quantity ELSE 0 END)) AS MPEGandAACDiferences
          FROM invoices i
          JOIN invoice_items ii ON i.InvoiceId = ii.InvoiceId
          JOIN tracks t ON ii.TrackId = t.TrackId
          JOIN media_types mt ON t.MediaTypeId = mt.MediaTypeId;"

result <- dbGetQuery(dbcon, query)
print(result)
```


## 17. Which country has the most customers?

```{r}
query <- "SELECT Country, COUNT(CustomerId) AS NumberOfCustomers
          FROM customers
          GROUP BY country
          ORDER BY NumberOfCustomers DESC
          LIMIT 1;"

result <- dbGetQuery(dbcon, query)
print(result)
```

## 18. How many tracks did "Amy Winehouse" record?

```{r}
query <- "SELECT COUNT(t.TrackId) AS NumberOfTracks
          FROM tracks t
          JOIN albums a ON t.AlbumId = a.AlbumId
          JOIN artists ar ON a.ArtistId = ar.ArtistId
          WHERE ar.Name = 'Amy Winehouse';"


result <- dbGetQuery(dbcon, query)
print(result)
```

## 19. What is the name of the artist who earned the most, i.e., whose aggregate total of all tracks sold is the highest?

```{r}

query <- "SELECT ar.Name AS ArtistName, SUM(ii.UnitPrice * ii.Quantity) AS TotalEarned
          FROM invoice_items ii
          JOIN tracks t ON ii.TrackId = t.TrackId
          JOIN albums a ON t.AlbumId = a.AlbumId
          JOIN artists ar ON a.ArtistId = ar.ArtistId
          GROUP BY ar.ArtistId, ar.Name
          ORDER BY TotalEarned DESC
          LIMIT 1;"

result <- dbGetQuery(dbcon, query)
print(result)
```


## 20. Which artist is the most popular in Paris? Popularity is defined in this context as the artist who was purchased most often by someone living in Paris.

```{r}
query <- "SELECT ar.Name AS ArtistName, COUNT(ii.TrackId) AS NumberOfPurchases
          FROM customers c
          JOIN invoices i ON c.CustomerId = i.CustomerId
          JOIN invoice_items ii ON i.InvoiceId = ii.InvoiceId
          JOIN tracks t ON ii.TrackId = t.TrackId
          JOIN albums a ON t.AlbumId = a.AlbumId
          JOIN artists ar ON a.ArtistId = ar.ArtistId
          WHERE c.City = 'Paris'
          GROUP BY ar.ArtistId
          ORDER BY NumberOfPurchases DESC
          LIMIT 1;"
result <- dbGetQuery(dbcon, query)
print(result)
```

